Proposed Solution

### 2017-04-16 OOPSLA17/secs/eval.tex
more consistent and thourough defs

\section{Implementation and Evaluation}

@@

We report the number of rules learned from the training set and the number of errors detected in the test set in Table~\ref{table:learning}.
One interesting note is that without probabilistic types we learned 327 fine grained rules and detected 1367 errors.
By introducing probabilistic types, we remove 114 rules and 1023 false positives.
We are guaranteed these are all false positives since there cannot be a rule of type $size*size=size$ because of the semantic interpretation of the $size$ units.

>>>>

\section{Implementation and Evaluation}
\label{sec:eval}

@@

We report the number of rules learned from the training set and the number of errors detected in the test set in Table~\ref{table:learning}.
One interesting note is that without probabilistic types we learned 327 fine grained rules and detected 1367 errors.
By introducing probabilistic types, we remove 114 incorrect rules and thereby remove 1023 false positives.
We are guaranteed these are all false positives since there cannot be a correct rule of relating the types $size*size$ and $size$ because of the semantic interpretation of the $size$ units.

### 2017-04-16 OOPSLA17/secs/learner.tex
rename fodler
more consistent and thourough defs

To learn rules over sets of configuration files, we use a generalization of \textit{association rule learning}~\cite{agrawal1993mining}, a technique that can be summarized as inductive machine learning.
Association rule learning is a technique to learn how frequently items of a set appear together.
For example, by examining a list of food store receipts, we would learn that when a customer buys bread and peanut butter, the set of purchased items is also likely to include jelly.
Since configuration files have complex relations, we extend these association relationships to generalized predicates.

A \textit{rule}, $r$, is then an implication relationship, $r = S \implies p(S,T)$, between two possibly empty sets of keywords $S,T \subset \keys$, which we call the source and target keyword sets.
The set $\keys$ is the set of unique keys from the training set (denoted \trainingSet) and the predicate $p$ is one of the classes of configuration errors.
Implicitly we interpret a rule to mean that if the keywords $S\cup T$ appear in a configuration file, the predicate $p$ should hold.
The task of the learning algorithm is to transform a training set to a set of rules, weighted with \textit{support} and \textit{confidence}.
The set of rules learned from training set \trainingSet constitutes a specification for a configuration file to be considered correct.

The two metrics, support and confidence, are used in association rule learning, as well as other rule based machine learning techniques~\cite{han2007frequent,langley1995applications}.
We use slightly modified definitions of of these to handle arbitrary predicates as rules.
During the learning process, each rule is assigned a support and confidence to measure the amount and quality of evidence for the rule.
%
\begin{equation*}
 \textit{support(r)} = \frac{|\{C \in \trainingSet \mid S_r \cup T_r \subseteq C\}|} {|\trainingSet|}
\end{equation*}
\begin{equation*}
 \textit{confidence(r)} = \frac{|\{C \in \trainingSet \mid p_r(S_r,T_r) \subseteq C\}| } {support(r)*|\trainingSet|}
\end{equation*}

Support is the frequency that the set of keywords in the proposed rule, $S \cup T$, have been seen in the configuration files $C$ in the training set \trainingSet.
%Notice support does not consider the predicate itself, only how many times the predicate had the chance to be evaluated.
Confidence is the percentage of times the rule predicate has held true over the given keywords.
In the learning process, each class of rule is manually assigned a support and confidence threshold, $t_s$ and $t_c$ respectively, below which a rule will be rejected for lack of evidence.
The denote the set rules that are learned and included as part of the final specification are as follows:

@@

\subsection{Error Classes}
Each class of error forms a rule with a predicate $p$, and additional restrictions on the sets $S, T$.
Ordering errors have $|S|, |T| = 1$ and use the predicate $order$ to mean the keyword $S$ must come before the keyword $T$ in any configuration file.
Missing keyword entry errors also require $|S|, |T| = 1$ and use the predicate $missing$ to mean the keyword $S$ must appear in the same file as the keyword $T$ in any configuration file.
The type rule is a set of rules over multiple predicates, which take the
form $S \implies \textit{isType}\ast(S)$, where $|S|=1, |T|=0$ and $\ast$ matches all the basic types (string, int, etc).

\app also supports two types of integer correlation rules, coarse-grained and fine-grained.
Both integer correlation rules are set of rules over the predicates $\{<,=,>\}$.
Coarse grain rules require $|S|, |T| = 1$, and the predicates have the typical interpretation.
Fine-grained rules use $|S|=2,|T|=1$, and interpret the predicates such that for $k_1,\ k_2 \in S,\ k1*k2$ must have the predicate relation to $T$.
The integer correlation rules also use probabilistic types to prune the search space.
To avoid learning too many false positives, we restrict this rule to either $size*int=size$, or $int*int = int$.
Without probabilistic typing, we would also learn, for example, $size*size=size$, which is an invalid interpretation.
%TODO include exactly how many false positive we prune with types over our training set? Here or in eval?

@@

With the rules generated by the learner module, \app checks whether any entry in a target configuration file violates the learned rules and constraints.
\app parses a verification target configuration file with the translator from Sec.~\ref{sec:trans}  to obtain a set of key-value pairs, $\mathcal{V}$, for that file.
Then, the checker applies the learner from Eq.~\ref{eq:learn},
$\textit{Learn}(\mathcal{V})$ to build the set of relations observed in the file, with the thresholds $t_s,t_c = 100\%$. 
The checker will then report the following set of errors:

\begin{align*}
\textit{Errors}(\mathcal{V}) = \{ r\ |\ & S_r \cup T_r \in \mathcal{V} \ \land \\
                               & r \in \textit{Learn}(\trainingSet) \ \land \\
                               & r \notin \textit{Learn}(\mathcal{V})
\end{align*}

>>>>

To learn rules over sets of configuration files, we use a generalization of \textit{association rule learning}~\cite{agrawal1993mining}, a technique very briefly summarized as inductive machine learning.
Association rule learning is used to learn how frequently items of a set appear together.
For example, by examining a list of food store receipts, we might learn that when a customer buys bread and peanut butter, the set of purchased items is also likely to include jelly.
Since configuration files have multiple complex relations, we extend these association relationships to generalized predicates.

A \textit{rule}, $r$, takes the form of an implication relationship, $r = S \in \validSet \implies T \in \validSet \land p(S,T)$, between two possibly empty sets of keywords $S,T \subset \keys$ where \validSet is the set of all valid configuration files.
In keeping with vocabulary from association rule learning, we refer to $S$ and $T$ as the source and target keyword sets respectively.
A rule states that if the source set of keyword appears in a valid configuration file, the target set of keywords must be present and the given predicate must hold.
For simplicity of notation, we treat the $S,T \in \validSet$ as implicit when writing a rule.

Taking the food store example, our learned rule would be $\{bread, butter\} \implies assoc(\{bread,butter\},\{jelly\})$.
The set $\keys$ is the set of unique keys from the training set (denoted \trainingSet) and the predicate $p$ is one of the classes of configuration errors.
The task of the learning algorithm is to transform a training set to a set of rules, weighted with \textit{support} and \textit{confidence}.
The set of rules learned from training set \trainingSet constitutes a specification for a configuration file to be considered correct.

The two metrics, support and confidence, are used in association rule learning, as well as other rule based machine learning techniques~\cite{han2007frequent,langley1995applications}.
We use slightly modified definitions of these, stated below, to handle arbitrary predicates as rules.
During the learning process, each rule is assigned a support and confidence to measure the amount and quality of evidence for the rule.
%
\begin{equation*}
 \textit{support(r)} = \frac{|\{C \in \trainingSet \mid S_r \cup T_r \subseteq C\}|} {|\trainingSet|}
\end{equation*}
\begin{equation*}
 \textit{confidence(r)} = \frac{|\{C \in \trainingSet \mid p_r(S_r,T_r) \subseteq C\}| } {support(r)*|\trainingSet|}
\end{equation*}

Support is the frequency that the set of keywords in the proposed rule, $S \cup T$, have been seen in the configuration files $C$ in the training set \trainingSet.
%Notice support does not consider the predicate itself, only how many times the predicate had the chance to be evaluated.
Confidence is the percentage of times the rule predicate has held true over the given keywords.
In the learning process, each class of rule is manually assigned a support and confidence threshold, $t_s$ and $t_c$ respectively, below which a rule will be rejected for lack of evidence.
We denote the set rules that are learned and included as part of the final specification are as follows:

@@

\subsection{Error Classes}
Each class of error forms a rule with a predicate $p$, and additional restrictions on the sets $S, T$.
Ordering errors require $|S|, |T| = 1$ and use the predicate $order$ to mean the keyword $S$ must come before the keyword $T$ in any configuration file.
For example, the ordering error given in Sec.~\ref{sec:motiv} is expressed:
%
\begin{equation*}
\{\texttt{innodb\_data\_home\_dir}\} \in \validSet \implies ord(\{\texttt{innoddb\_data\_home\_dir}\},\{\texttt{innodb\_data\_file\_path}\})
\end{equation*}
%
Missing keyword entry errors require $|S|, |T| = 1$ and use the predicate $missing$ to mean the keyword $T$ must appear in the same file as the keyword $S$ in any configuration file.
The type rule is a set of rules over multiple predicates, where $|S|=1, |T|=0$.
These type rules take the form $S \implies is\_(S)$ where $\_$ matches all the basic types (bool, int, size, etc).

\app also supports two types of integer correlation rules, coarse-grained and fine-grained.
Both integer correlation rules are set of rules over the predicates $\{<,=,>\}$.
Coarse grain rules require $|S|, |T| = 1$, and the predicates have the typical interpretation.
Fine-grained rules use $|S|=2,|T|=1$, and interpret the predicates such that for $k_1,\ k_2 \in S,\ k1*k2$ must have the predicate relation to $T$.
The integer correlation rules also use probabilistic types to prune the search space.
To avoid learning too many false positives, we restrict this rule to either $size*int=size$, or $int*int = int$, as shown in Fig.~\ref{fig:allrules}.
Without probabilistic typing, we would also learn rules which do not have a invalid sematinc interpretation, for example a relation between three size keywords ($k1,k2,k3::size$) of the form $k1*k2>k3$.

{
\setlength{\abovecaptionskip}{-.05pt}
\setlength{\belowcaptionskip}{-15pt}
\begin{figure}
\begin{mathpar}
%\hspace{-1cm} %no idea why, but this is better than \centering here
\inferrule* [Right=bool]
{k1 :: bool}
{isbool(k1) :: Rule}
\and\quad
\inferrule* [Right=int]
{k1 :: int}
{isint(k1) :: Rule}
\and\\
\inferrule* [Right=fine\_grain\_rule]
{k1,k2,k3 :: int}
{fine(k1,k2,k3) :: Rule}
\and\quad
\inferrule* [Right=fine\_grain\_rule]
{k1,k3 :: size \\ k2 :: int}
{fine(k1,k2,k3) :: Rule}
\end{mathpar}
\caption{An expanded set of typing judgements for valid rules}
\label{fig:allrules}
\end{figure}
}

@@

With the rules generated by the learner module, \app checks whether any entry in a target configuration file violates the learned rules and constraints.
\app parses a single verification target configuration file with the translator from Sec.~\ref{sec:trans}  to obtain a set of key-value pairs, $\mathcal{C}$, for that file.
Then, the checker applies the learner from Eq.~\ref{eq:learn} with $\textit{Learn}(\mathcal{C})$,
  to build the set of relations observed in the file, with the thresholds $t_s,t_c = 100\%$. 
The checker will then report the following set of errors:

\begin{align*}
\textit{Errors}(\mathcal{C}) = \{ r\ |\ & S_r \cup T_r \in \mathcal{C} \ \land \\
                               & r \in \textit{Learn}(\trainingSet) \ \land \\
                               & r \notin \textit{Learn}(\mathcal{C})
\end{align*}

### 2017-04-16 OOPSLA17/secs/postanal.tex
rename fodler
more consistent and thourough defs

The learner outputs a set of rules learned from the training set as
described in Sec.~\ref{sec-learn}. Recall that a rule is 
an implication relationship of the form $r = S \implies p(S,T)$.
This data is necessary to perform the core verification task,
but can also be used for further analysis. By interpreting the rules as a graph (which
we call the \textit{rule graph}), we can 
use tools from graph theory to extract information
about the configuration space that can improve the quality of the learned model.
We inspect properties of this 
rule graph to sort reported errors by those most likely to be valid.
To demonstrate the additional value of the rule graph, we
also use it to estimate the complexity of a configuration file.

Accessibility of the rule graph is a useful property of the association 
rule learning technique applied by \app.
While it is possible to analyze the models from other machine learning techniques, such as neural networks~\cite{lei2016rationalizing} and
conditional random fields~\cite{raychev2015predicting}, these analyses require a deep knowledge of the applied techniques.
In contrast the rule graph is a relatively simple, yet information rich, representation of the learned model.
The following section provides a precise definition of the rule graph
and demonstrates useful metrics we derive for the purposes of ranking reported errors
and complexity analysis.

@@

The results from ranking errors in this way are presented later in
the paper.

>>>>

The learner outputs a set of rules learned from the training set as
described in Sec.~\ref{sec-learn}. Recall that a rule is 
an implication relationship of the basic form $r = S \implies p(S,T)$.
This data is necessary to perform the core verification task,
but can also be used for further analysis. By interpreting the rules as a graph (which
we call the \textit{rule graph}), we can 
use tools from graph theory to extract information
about the configuration space that can improve the quality of the learned model.
We inspect properties of this 
rule graph to sort reported errors by those most likely to be valid.
To demonstrate the additional value of the rule graph, we
also use it to estimate the complexity of a configuration file.

Accessibility of the rule graph is a useful property of the association 
rule learning technique applied by \app.
While it is possible to analyze the models from other machine learning techniques, such as neural networks~\cite{lei2016rationalizing} and
conditional random fields~\cite{raychev2015predicting}, these analyses require a deep knowledge of the applied techniques.
In contrast, the rule graph is a relatively simple, yet information rich, representation of the learned model.
The following section provides a precise definition of the rule graph
and demonstrates useful metrics we derive for the purposes of ranking reported errors
and complexity analysis.

@@

The results from ranking errors in this way are presented in Sec.~\ref{sec:eval}.

### 2017-04-16 OOPSLA17/secs/related.tex
more consistent and thourough defs

>>>>

\para{Association Rule Learning}
This approach is not only generalizes association rule learning, but also another popular learning strategy called \textit{sequential pattern mining}~\cite{seqMine}.
Our ordering rules are similar to the rules learned in sequential pattern mining, although can restrict our ordering rules to |S|,|T|=1 since these are the most common misconfigurations we encounter in practice.
While a major limitation to sequential pattern mining is the scalability of the problem, we escape this issue since configuration files tend to be no more than thousands of lines at a time.

### 2017-04-16 OOPSLA17/secs/trans.tex
more consistent and thourough defs

\para{Probabilistic Types}
%\label{sec:ptypes}
An additional challenge is that it is not always possible to fully determine the type of key based on a single example value. 
For this reason, we introduce \textit{probabilistic types}, as contrasted with \textit{basic types}.
In \app, the set of basic types contains strings, file paths, integers, sizes, and Booleans. 
Taking the configuration \ref{lst:file1.cnf}, we can assume {\tt foo} is a Boolean type by the grammar of MySQL,
  but the keyword {\tt bar} could be many types.
If we choose the type based on the first example, {\tt bar} will be a integer type.
If we choose a type that fits all examples, {\tt bar} will be a string.
However the correct classification needed is a Boolean type.
Let us assume that there is a critical rule we must learn that the Boolean keywords should have the same values, $eq(\texttt{foo},\texttt{bar})$.
If we take {\tt bar}$::int$, we do not learn the above rule, nor do we learn this rule with {\tt bar}$::string$ - only with {\tt bar}$::bool$ is the rule is valid.
To resolve this ambiguity, and choose the best type, the translator assigns a distribution of types to a keyword based on examples from the training set of configuration files (denoted $\trainingSet=\{C\}$).

@@

{
\setlength{\abovecaptionskip}{-.05pt}
\setlength{\belowcaptionskip}{-15pt}
\begin{figure}
\begin{mathpar}
\hspace{-1cm} %no idea why, but this is better than \centering here
\inferrule* [Right=ptype]
{c_{int} = |\{\forall C \in \trainingSet.\ \forall (k,v) \in C.\ v \in \mathbb{Z}\}| \\ c_{bool} = |\{\forall C \in \trainingSet.\ \forall (k,v) \in C.\ v \in \{0,1,ON,OFF\}\}| }
{k : \ptype[int = c_{int}, bool = c_{bool}]}
\and\\
\inferrule* [Right=int]
{k:\ptype \\ p_{int}(\ptype)}
{k :: int}
\and
\inferrule* [Right=bool]
{k:\ptype \\ p_{bool}(\ptype)}
{k :: bool}
\and
\inferrule* [Right=eq\_rule]
{k_1 :: \tau \\ k_2 :: \tau}
{eq(k_1,k_2) :: Rule}
\end{mathpar}
\caption{Type judgments for a probabilistic type system with $\mathcal{T} = \{bool,int\}$ and an equality rule}
\label{fig:ptypes}
\end{figure}
}

@@

To resolve the ambiguity in this case, we must add a new rule {\scriptsize INTBOOL} that introduces a new, internal type $intbool$ to be created when both resolution predicates are true.
The antecedent statement of {\scriptsize INT} must also be updated to $p_{int}(\ptype) \land \neg p_{bool}(\ptype)$, and likewise with {\scriptsize BOOL}.
The $intbool$ type is only a placeholder to be used with the additional subtyping relations $intbool <: int$ and $intbool <: bool$
These subtype relations allow the $intbool$ to take the place of either $int$ or $bool$ when determining if two keywords may be compared in the {\scriptsize EQ\_RULE}.

\begin{mathpar}
\hspace{-0.8cm} %no idea why, but this is better than \centering here
\inferrule* [Right=intbool]
{k:\ptype \\ p_{bool}(\ptype) \\ p_{int}(\ptype)}
{k :: intbool}
\and
\inferrule* [Right=order\_rule]
{k1,k2 \in C \\ k1 \neq k2}
{ord(k1,k2) :: Rule}
\end{mathpar}

>>>>

\para{Probabilistic Types}
%\label{sec:ptypes}
An additional challenge is that it is not always possible to fully determine the type of key based on a single example value. 
For this reason, we introduce \textit{probabilistic types}, as contrasted with \textit{basic types}.
In \app, the set of basic types contains strings, file paths, integers, sizes, and Booleans. 
Taking the configuration \ref{lst:file1.cnf}, we can assume {\tt foo} is a Boolean type by the grammar of MySQL,
  but the keyword {\tt bar} could be many types.
If we choose the type based on the first example, {\tt bar} will be a integer type.
If we choose a type that fits all examples, {\tt bar} will be a string.
However the correct classification needed is a Boolean type.
For illustrative purposes, let us assume that there is a critical rule we must learn where tow Boolean keywords should always have the same values, e.g. $eq(\texttt{foo},\texttt{bar})$.
If we take {\tt bar}$::int$, we do not learn the above rule, nor do we learn this rule with {\tt bar}$::string$ - only with {\tt bar}$::bool$ is the rule is valid.
To resolve this ambiguity, and choose the best type, the translator assigns a distribution of types to a keyword based on examples from the training set of configuration files (denoted $\trainingSet=\{C\}$).

@@

{
\setlength{\abovecaptionskip}{-.05pt}
\setlength{\belowcaptionskip}{-15pt}
\begin{figure}
\begin{mathpar}
%\hspace{-1cm} %no idea why, but this is better than \centering here
\inferrule* [Right=ptype]
{c_{int} = |\{\forall C \in \trainingSet.\ \forall (k,v) \in C.\ v \in \mathbb{Z}\}| \\\\ c_{bool} = |\{\forall C \in \trainingSet.\ \forall (k,v) \in C.\ v \in \{0,1,ON,OFF\}\}| }
{k : \ptype[int = c_{int}, bool = c_{bool}]}
\and\\
\inferrule* [Right=int]
{k:\ptype \\ p_{int}(\ptype)}
{k :: int}
\and
\inferrule* [Right=bool]
{k:\ptype \\ p_{bool}(\ptype)}
{k :: bool}
\and\\
\inferrule* [Right=eq\_rule]
{k_1 :: \tau \\ k_2 :: \tau}
{eq(k_1,k_2) :: Rule}
\and
\inferrule* [Right=order\_rule]
{k1,k2 \in C \\ k1 \neq k2}
{ord(k1,k2) :: Rule}
\end{mathpar}
\caption{Type judgments for a probabilistic type system with $\mathcal{T} = \{bool,int\}$ and an equality and ordering rule}
\label{fig:ptypes}
\end{figure}
}

@@
